Abundant base ratio power analysis
========================================================
### Objective:  
Perform power analysis to determine the desired coverage for estimating abundant base ratios.

#### Assumptions:  
* The total gene copy number for _E. coli_ and _L. monocytogenes_ is 7 and 6 respectively.
* Sequnced bases follow a binomial distribution

```{r}
library(reshape2)
library(ggplot2)
library(stringr)
library(xtable)
sessionInfo()
```

### Simulated data for power analysis
```{r}
coverage <- c(1:500,seq(from = 510, to = 10000, by = 10))
ratio_levels <- c("x6:1","x5:1","x4:3","x3:3")
power_df <- data.frame(coverage)
ratios <-c(6/7,5/6,4/7,3/6)
for(i in 1:4){
  power_cal <- sqrt((ratios[i]*(1-ratios[i])/coverage))
  power_df[,ratio_levels[i]] <- power_cal                  
}
power_dfm <- melt(power_df, id = "coverage")
power_dfm$Ratio <- str_replace(power_dfm$variable, "x","")
```

### Theoretical decay plot assuming a binomial distribution
```{r fig.width=7, fig.height=6}
ggplot(power_dfm) + geom_line(aes(x = coverage, y = value, color = Ratio)) + scale_x_log10() + labs(x = "Desired Coverage", y = "Proportion Standard Deviation") + theme_bw()
```


### Power analysis for consensus estimated abundant base ratios
Desired standard deviation is defined as 1/4 of 1/gene copy number.  This is based on a 95 % confidence interval of +/- 2 x standard deviation  and no overlap between the confidence invervals for the possible abundant base copy ratios, e.g. for _L. monocyotogenes_ 1/6, 2/6, ...6/6.
```{r}
pwr <- data.frame(abd_ratio = c("6:1","5:1","4:3","3:3"), ratios, sd = 1/(4*(c(7,6,7,6))))
pwr$cov <- (pwr$ratios * (1-pwr$ratios))/pwr$sd^2
colnames(pwr) <- c("Abundant Base Ratios","Abundant Base Proportions","Standard Deviation", "Desired Coverage")                                  
```

```{r, results='asis', echo=FALSE}
coverage_table <- xtable(pwr, caption = "Desired coverage for consensus abundant base ratios",
                         align = c("c","c","c","c","c"), digits = 3)
print.xtable(coverage_table, type = "html", caption.placement = "top", include.rownames=FALSE)
```

### Function from Steve
```{r}
pwr<-function(n.copy, prob=.95){
  p<-(1:(n.copy-1))/n.copy
  N<-p
  for(i in 1:length(p)){
    n<-0;t.prob<-0
    while(t.prob<prob){
      n<-n+20
      t.prob<-pbinom(min(round((p[i]+.5/n.copy)*n),n),n,p[i])-pbinom(max(round((p[i]-.5/n.copy)*n)-1,0),n,p[i])
    }
    while(t.prob>prob){
      n<-n-1
      t.prob<-pbinom(min(round((p[i]+.5/n.copy)*n),n),n,p[i])-pbinom(max(round((p[i]-.5/n.copy)*n)-1,0),n,p[i])
    }
    N[i]<-n+1
  }
  names(N)<-paste(1:(n.copy-1),(n.copy-1):1,sep=":")
  N
}
```

